---
title: "NORTHSIDE ISD HIGH SCHOOL GRADUATES TO LOCAL HIGHER ED"
output: 
  flexdashboard::flex_dashboard
---

Page 1: Introduction  {data-orientation=columns}
===================================== 
Column {data-width=400}
-------------------------------------  
James Ford  
STA 6233: Advanced R Programming  
Spring 2021   
<hr>

### **Background**

This report analyzes the high school graduates from Northside ISD (NISD) in San Antonio and the extent to which they enrolled in college locally.  

Data was downloaded from the Texas Higher Education Coordinating Board (THECB) for Class of 2017 - 2019 high school graduates, as well as 6-year college degree completion rates.  

The following pages detail the data results.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, echo = FALSE, warning = FALSE, error = FALSE, include=FALSE}
library(leaflet)
library(tigris)


# Make Icons
hs <- makeIcon(iconUrl = "https://github.com/JamieFord/STA-6233-Advanced-R-Programming/blob/d6d1918ce222614adb3d04a9051e49b21d21b422/school-icon.png?raw=true",
               iconWidth = 15, iconHeight = 15,
               iconAnchorX = 9, iconAnchorY = 9,
               #shadowUrl = "http://leafletjs.com/examples/custom-icons/leaf-shadow.png",
               #shadowWidth = 50, shadowHeight = 64,
               #shadowAnchorX = 4, shadowAnchorY = 62
)

college <- makeIcon(iconUrl = "https://github.com/JamieFord/STA-6233-Advanced-R-Programming/blob/d6d1918ce222614adb3d04a9051e49b21d21b422/university-icon.png?raw=true",
                    iconWidth = 20, iconHeight = 20,
                    iconAnchorX = 9, iconAnchorY = 9,
                    #shadowUrl = "http://leafletjs.com/examples/custom-icons/leaf-shadow.png",
                    #shadowWidth = 50, shadowHeight = 64,
                    #shadowAnchorX = 4, shadowAnchorY = 62
)
```

```{r echo = FALSE, warning = FALSE, error = FALSE, include=FALSE}
# Place Icons on Map
school <- school_districts("TX", cb=TRUE) %>% 
  leaflet() %>% 
  #addTiles() %>% 
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(popup=~NAME,
              fillColor = "white",
              color = "black",
              weight = 1,
              highlight = highlightOptions(
                weight = 3,
                fillOpacity = 0,
                color = "black",
                opacity = 1.0,
                bringToFront = TRUE,
                sendToBack = TRUE) ) %>%
  setView(lng=-98.6015932, lat=29.4905564, zoom = 10) %>%
  addMarkers(lng=-98.6189, lat=29.5827, icon = college,
             label ="UTSA", popup="<a href='https://www.utsa.edu'>UTSA</a>") %>%
  addMarkers(lng=-98.704280577599, lat=29.472074401384, icon = college, label ="NW Vista College", popup="<a href='https://www.alamo.edu/nvc/'>NWV</a>") %>%
  addMarkers(lng=-98.4943314, lat=29.446188, icon = college, label ="San Antonio College (SAC)", popup="<a href='https://www.alamo.edu/sac/'>SAC</a>") %>%
  addMarkers(lng=-98.546252491408, lat=29.322890202361, icon = college, label ="Palo Alto College", popup="<a href='https://www.alamo.edu/pac/'>PAC</a>") %>%
  addMarkers(lng=-98.453940998101, lat=29.41544254087, icon = college, label ="St. Philips College", popup="<a href='https://www.alamo.edu/spc/'>SPC</a>") %>%
  addMarkers(lng=-98.5434, lat=29.3294, icon = college, label ="TAMU-SA", popup="<a href='https://www.tamusa.edu/'>TAMUSA</a>") %>%
  addMarkers(lng=-98.562637, lat=29.453159, icon = college, label ="St. Mary's", popup="<a href='https://www.stmarytx.edu/'>StMU</a>") %>%
  addMarkers(lng=-98.472252, lat=29.470289, icon = college, label ="Incarnate Word", popup="<a href='https://www.uiw.edu/'>IWC</a>") %>%
  addMarkers(lng=-98.544159, lat=29.426729, icon = college, label ="OLLU", popup="<a href='https://www.ollusa.edu/'>OLLU</a>") %>%
  addMarkers(lng=-98.617371, lat=29.46546, icon = hs, label ="Holmes HS", popup="<a href='https://www.nisd.net/holmes'>Holmes</a>") %>%
  addMarkers(lng=-98.640387024304, lat=29.41885075216, icon = hs, label ="Jay HS", popup="<a href='https://www.nisd.net/jay'>Jay</a>") %>%
  addMarkers(lng=-98.5837253, lat=29.5655711, icon = hs, label ="Clark HS", popup="<a href='https://www.nisd.net/clark'>Clark</a>") %>%
  addMarkers(lng=-98.6202986, lat=29.5135634, icon = hs, label ="Marshall HS", popup="<a href='https://www.nisd.net/marshall'>Marshall</a>") %>%
  addMarkers(lng=-98.720273, lat=29.500035, icon = hs, label ="Taft HS", popup="<a href='https://www.nisd.net/taft'>Taft</a>") %>%
  addMarkers(lng=-98.6856206883, lat=29.5590895, icon = hs, label ="O'Connor HS", popup="<a href='https://www.nisd.net/oconnor'>O'Connor</a>") %>%
  addMarkers(lng=-98.5752975 , lat=29.5216186, icon = hs, label ="Health Careers HS", popup="<a href='https://www.nisd.net/healthcareers'>HCHS</a>") %>%
  addMarkers(lng=-98.7246428 , lat=29.457912, icon = hs, label ="Brennan HS", popup="<a href='https://www.nisd.net/brennan'>Brennan</a>") %>%
  addMarkers(lng=-98.646286, lat=29.565452, icon = hs, label ="Brandeis HS", popup="<a href='https://www.nisd.net/brandeis'>Brandeis</a>") %>%
  addMarkers(lng=-98.684628, lat=29.439971, icon = hs, label ="Stevens HS", popup="<a href='https://www.nisd.net/stevens'>Stevens</a>") %>%
  addMarkers(lng=-98.770592, lat=29.509619, icon = hs, label ="Harlan HS", popup="<a href='https://www.nisd.net/harlan'>Harlan</a>") %>%
  addMarkers(lng=-98.669920, lat=29.463730, icon = hs, label ="Warren HS", popup="<a href='https://www.nisd.net/warren'>Warren</a>") %>%
  addMarkers(lng=-98.586130, lat=29.554580, icon = hs, label ="Chavez Excel Acad", popup="<a href='https://www.nisd.net/excel'>Excel</a>")
```

Column {data-width=600}
-------------------------------------
### **Map:** *Zoom in, mouse over, and click on icons (HS and College Campuses) and polygons (School Districts)*

```{r echo = FALSE, warning = FALSE, error = FALSE}
school
```

Page 2: NISD High Schools {data-orientation=columns}
===================================== 

Column {data-width=400}
-------------------------------------

```{r, echo = FALSE, warning = FALSE, error = FALSE, include=FALSE}

library(readxl)
library(dplyr)
library(tidyr)
library(ggplot2)
library(RColorBrewer)
library(scales)
library(plotly)
library(knitr)
library(kableExtra)
library(parcats)
library(easyalluvial)

op <- options(gvis.plot.tag = "chart")

# Read 2019 Master Enroll file
HSGrads2019 <- read_excel("HSGradsEnrolledCampus2019(1).xlsx", col_types = c("text", "text", "text", "text", "text", "numeric"))
summary(HSGrads2019)

# Total Grads Count
TotGrads <- HSGrads2019 %>% 
  filter(Institution=='Total high school graduates')

#No College
NoCollege <- HSGrads2019 %>% 
  filter(Institution %in% c('Not trackable','Not found')) %>%
  group_by(Name) %>%
  summarize(Students=sum(Students))

#Split into College/Univ dfs
UTSA <- HSGrads2019 %>% 
  filter(Institution=='U. OF TEXAS AT SAN ANTONIO (010115)')

NWV <- HSGrads2019 %>% 
  filter(Institution=='ALAMO CCD NW VISTA COLLEGE (000307)')

SAC <- HSGrads2019 %>% 
  filter(Institution=='ALAMO CCD SAN ANTONIO COLLEGE (009163)')

PAC <- HSGrads2019 %>% 
  filter(Institution=='ALAMO CCD PALO ALTO COLLEGE (023413)')

SPC <- HSGrads2019 %>% 
  filter(Institution=='ALAMO CCD ST. PHILIPS COLLEGE (003608)')

TAMUSA <- HSGrads2019 %>% 
  filter(Institution=='TEXAS A&M UNIV-SAN ANTONIO (042485)')

IWC <- HSGrads2019 %>% 
  filter(Institution=='UNIV OF THE INCARNATE WORD (003578)')

StMU <- HSGrads2019 %>% 
  filter(Institution=="ST. MARY'S UNIVERSITY (003623)")

OLLU <- HSGrads2019 %>% 
  filter(Institution=="OUR LADY OF THE LAKE UNIV/SA (003598)")

District2019 <- data_frame(College=c("UTSA","NWV","PAC","SAC","SPC","TAMUSA","IWC","StMU","OLLU"),
                           Year='2019',
                           Enroll=c(sum(UTSA$Students),sum(NWV$Students),sum(PAC$Students),sum(SAC$Students),
                                    sum(SPC$Students),sum(TAMUSA$Students),sum(IWC$Students),
                                    sum(StMU$Students),sum(OLLU$Students)))

# Merge into HS dfs
HS2019 <- cbind.data.frame(UTSA$Name,UTSA$Students) %>%
  rename(Name=1,UTSA=2)

HS2019 <- left_join(HS2019,NWV) %>%
  select(-County,-Code,-Institution,-District) %>%
  rename(NWV=Students)

HS2019 <- left_join(HS2019,PAC) %>%
  select(-County,-Code,-Institution,-District) %>%
  rename(PAC=Students)

HS2019 <- left_join(HS2019,SAC) %>%
  select(-County,-Code,-Institution,-District) %>%
  rename(SAC=Students)

HS2019 <- left_join(HS2019,SPC) %>%
  select(-County,-Code,-Institution,-District) %>%
  rename(SPC=Students)

HS2019 <- left_join(HS2019,TAMUSA) %>%
  select(-County,-Code,-Institution,-District) %>%
  rename(TAMUSA=Students)

HS2019 <- left_join(HS2019,IWC) %>%
  select(-County,-Code,-Institution,-District) %>%
  rename(IWC=Students)

HS2019 <- left_join(HS2019,StMU) %>%
  select(-County,-Code,-Institution,-District) %>%
  rename(StMU=Students)

HS2019 <- left_join(HS2019,OLLU) %>%
  select(-County,-Code,-Institution,-District) %>%
  rename(OLLU=Students)

HS2019 <- left_join(HS2019,TotGrads) %>%
  select(-County,-Code,-Institution,-District) %>%
  rename(TotGrads=Students)

# Calc Percentages by College/Univ
HS2019$UTSA_Pct <- round(HS2019$UTSA / HS2019$TotGrads * 100, 1)
HS2019$NWV_Pct <- round(HS2019$NWV / HS2019$TotGrads * 100, 1)  
HS2019$PAC_Pct <- round(HS2019$PAC / HS2019$TotGrads * 100, 1)  
HS2019$SAC_Pct <- round(HS2019$SAC / HS2019$TotGrads * 100, 1) 
HS2019$SPC_Pct <- round(HS2019$SPC / HS2019$TotGrads * 100, 1) 
HS2019$TAMUSA_Pct <- round(HS2019$TAMUSA / HS2019$TotGrads * 100 , 1) 
HS2019$IWC_Pct <- round(HS2019$IWC / HS2019$TotGrads * 100 , 1)
HS2019$StMU_Pct <- round(HS2019$StMU / HS2019$TotGrads * 100 , 1)
HS2019$OLLU_Pct <- round(HS2019$OLLU / HS2019$TotGrads * 100 , 1)

HS_piv2019 <- HS2019 %>%
  pivot_longer(cols = UTSA_Pct:OLLU_Pct, names_to = "College", values_to = "Pct")

# Graph
mycolors = c(brewer.pal(name="Accent", n = 3), brewer.pal(name="Dark2", n = 6))
Bar_Graph1 <- ggplot(data=HS_piv2019, aes(x = reorder(Name, desc(Name)), y=Pct, fill=College)) +
  geom_bar(stat="identity") +
  coord_flip() + 
  theme_classic() +
  theme(axis.text.y = element_text(face="bold")) + 
  scale_fill_manual(values = mycolors) + 
  #scale_y_continuous(labels = scales::percent) +
  labs(y="All Graduates (%)") +
  theme(axis.title.y = element_blank()) +
  theme(legend.title = element_blank())

# Table
Table1_df <- HS2019 %>%
  select(1:11) %>%
  select("Name", sort(colnames(.))) %>%
  rename(HS="Name") %>%
  relocate(TotGrads, .after = HS)

# SANKEY Graph
HS2019_sk3 <- left_join(HS2019,NoCollege) %>%
  rename(NoCollege=Students)
HS2019_sk3 <- HS2019_sk3 %>%
  mutate("LocalCollege" = rowSums((HS2019_sk3[,2:10]), na.rm = TRUE)) %>%
  mutate("NonLocal" = TotGrads-LocalCollege-NoCollege)

HS_piv2019_sk3 <- HS2019_sk3 %>%
  pivot_longer(cols = NoCollege:NonLocal, names_to = "Path", values_to = "Grads")


HS2019_Sk <- data.frame(From=HS_piv2019_sk3$Name, 
                  To=HS_piv2019_sk3$Path, 
                  Weight=HS_piv2019_sk3$Grads)

HS_test <-  HS2019_Sk[rep(1:nrow(HS2019_Sk), HS2019_Sk$Weight),]

```

### **HS Graduates**  
A total of `r sum(TotGrads$Students)` students graduated from NISD high schools in the Class of 2019 (the latest data available).  

The interactive graphs/table to the right depict: (1) the number and proportion of graduates by high school and whether they enrolled locally in higher education institutions, non-local institutions, or did not enroll in college; (2) the percent share each local college received from each high school; and, (3) actual counts of HS graduates enrolling locally.

*Note that Trinity University was not listed in the THECB figures as the enrollment counts per high school were below the reporting threshold, and are therefore included with non-local colleges.*  

Key findings include:  
$*$ A total of `r round(sum(HS2019_sk3$LocalCollege)/sum(HS2019_sk3$TotGrads)*100,0)`% of 2019 NISD HS graduates enrolled in local higher education institutions.  
$*$ Health Careers HS had the highest percentage of graduates both enrolling in any local institution as well as the most choosing UTSA.  
$*$ Except for Health Careers, Northwest Vista College (NWV) was the primary institution chosen by NISD graduates.  


Column {.tabset}
-------------------------------------
### Graph: Grads by Enroll
```{r echo = FALSE, warning = FALSE, error = FALSE}
p = alluvial_wide( select(HS_test, From, To)
                              , fill_by = 'first_variable' )

parcats(p, marginal_histograms = FALSE, data_input = HS_test,hoveron = "category", width = "60%")

```


### Graph: Enroll Local College

```{r echo = FALSE, warning = FALSE, error = FALSE}
ggplotly(Bar_Graph1,tooltip = "y")
```

### Table: Enroll Local College

```{r echo = FALSE, warning = FALSE, error = FALSE}
Table1 <-  kbl(Table1_df) %>%
  kable_styling("hover")
Table1
```

Page 3: Local Higher Ed {data-orientation=columns}
===================================== 

```{r, echo = FALSE, warning = FALSE, error = FALSE, include=FALSE}

# Read 2018 Master Enroll file
HSGrads2018 <- read_excel("HSGradsEnrolledCampus2018.xlsx", col_types = c("text", "text", "text", "text", "text", "numeric"))
summary(HSGrads2018)

# Total Grads Count
TotGrads <- HSGrads2018 %>% 
  filter(Institution=='Total high school graduates')

#Split into College/Univ dfs
UTSA <- HSGrads2018 %>% 
  filter(Institution=='U. OF TEXAS AT SAN ANTONIO (010115)')

NWV <- HSGrads2018 %>% 
  filter(Institution=='ALAMO CCD NW VISTA COLLEGE (000307)')

SAC <- HSGrads2018 %>% 
  filter(Institution=='ALAMO CCD SAN ANTONIO COLLEGE (009163)')

PAC <- HSGrads2018 %>% 
  filter(Institution=='ALAMO CCD PALO ALTO COLLEGE (023413)')

SPC <- HSGrads2018 %>% 
  filter(Institution=='ALAMO CCD ST. PHILIPS COLLEGE (003608)')

TAMUSA <- HSGrads2018 %>% 
  filter(Institution=='TEXAS A&M UNIV-SAN ANTONIO (042485)')

IWC <- HSGrads2018 %>% 
  filter(Institution=='UNIV OF THE INCARNATE WORD (003578)')

StMU <- HSGrads2018 %>% 
  filter(Institution=="ST. MARY'S UNIVERSITY (003623)")

OLLU <- HSGrads2018 %>% 
  filter(Institution=="OUR LADY OF THE LAKE UNIV/SA (003598)")

District2018 <- data_frame(College=c("UTSA","NWV","PAC","SAC","SPC","TAMUSA","IWC","StMU","OLLU"),
                           Year='2018',
                           Enroll=c(sum(UTSA$Students),sum(NWV$Students),sum(PAC$Students),sum(SAC$Students),
                                    sum(SPC$Students),sum(TAMUSA$Students),sum(IWC$Students),
                                    sum(StMU$Students),sum(OLLU$Students)))

# Read 2017 Master Enroll file
HSGrads2017 <- read_excel("HSGradsEnrolledCampus2017.xlsx", col_types = c("text", "text", "text", "text", "text", "numeric"))
summary(HSGrads2018)

# Total Grads Count
TotGrads <- HSGrads2017 %>% 
  filter(Institution=='Total high school graduates')

#Split into College/Univ dfs
UTSA <- HSGrads2017 %>% 
  filter(Institution=='U. OF TEXAS AT SAN ANTONIO (010115)')

NWV <- HSGrads2017 %>% 
  filter(Institution=='ALAMO CCD NW VISTA COLLEGE (000307)')

SAC <- HSGrads2017 %>% 
  filter(Institution=='ALAMO CCD SAN ANTONIO COLLEGE (009163)')

PAC <- HSGrads2017 %>% 
  filter(Institution=='ALAMO CCD PALO ALTO COLLEGE (023413)')

SPC <- HSGrads2017 %>% 
  filter(Institution=='ALAMO CCD ST. PHILIPS COLLEGE (003608)')

TAMUSA <- HSGrads2017 %>% 
  filter(Institution=='TEXAS A&M UNIV-SAN ANTONIO (042485)')

TAMUSA <- HSGrads2017 %>% 
  filter(Institution=='TEXAS A&M UNIV-SAN ANTONIO (042485)')

IWC <- HSGrads2017 %>% 
  filter(Institution=='UNIV OF THE INCARNATE WORD (003578)')

StMU <- HSGrads2017 %>% 
  filter(Institution=="ST. MARY'S UNIVERSITY (003623)")

OLLU <- HSGrads2017 %>% 
  filter(Institution=="OUR LADY OF THE LAKE UNIV/SA (003598)")

District2017 <- data_frame(College=c("UTSA","NWV","PAC","SAC","SPC","TAMUSA","IWC","StMU","OLLU"),
                           Year='2017',
                           Enroll=c(sum(UTSA$Students),sum(NWV$Students),sum(PAC$Students),sum(SAC$Students),
                                    sum(SPC$Students),sum(TAMUSA$Students),sum(IWC$Students),
                                    sum(StMU$Students),sum(OLLU$Students)))

#Data for Line Graph
Colleges <- data.frame(rbind(District2017,District2018,District2019))

Line_Graph1 <- ggplot(Colleges, aes(x=Year, y=Enroll, group=College)) +
  geom_line(aes(color=College)) +
  geom_point(aes(color=College)) + 
  theme_classic() +
  scale_color_manual(values = mycolors) +
  theme(legend.title = element_blank()) + 
  scale_y_continuous(limits = c(0, 1500))

Tot_Colleges_2019 <- Colleges %>% filter(Year == 2019)
Tot_Colleges_2017 <- Colleges %>% filter(Year == 2017)
Pct_Change <- round((sum(Tot_Colleges_2019$Enroll) - sum(Tot_Colleges_2017$Enroll)) / sum(Tot_Colleges_2017$Enroll) * 100, 0)


```




```{r, echo = FALSE, warning = FALSE, error = FALSE, include=FALSE}
# Outcomes
HSGradsOutcomes <- read_excel("HSGradsOutcomes.xlsx", col_types = c("skip", "text", "skip","text","skip",                                      "text", "numeric","numeric",  "skip", "numeric", "numeric","numeric"))
summary(HSGradsOutcomes)

HSGradsOutcomes$Degree <- HSGradsOutcomes$`Earned Assoc.` + HSGradsOutcomes$`Earned Bac. or Higher`

Outcome <- HSGradsOutcomes %>% 
  filter(`Bachelor's Degree Granting Institution and College Graduation Status` %in% 
           c('U. OF TEXAS AT SAN ANTONIO','ALAMO CCD NW VISTA COLLEGE','ALAMO CCD PALO ALTO COLLEGE',
             'ALAMO CCD SAN ANTONIO COLLEGE','ALAMO CCD ST. PHILIPS COLLEGE','OUR LADY OF THE LAKE UNIV/SA',
             "ST. MARY'S UNIVERSITY",'UNIV OF THE INCARNATE WORD','TEXAS A&M UNIV-SAN ANTONIO')) %>%
  rename(College=`Bachelor's Degree Granting Institution and College Graduation Status`) %>%
  group_by(College) %>%
  summarize(Grads_From_NISD=sum(Degree)) %>%
  mutate(College=recode(College,
                        'U. OF TEXAS AT SAN ANTONIO'="UTSA",
                        'ALAMO CCD NW VISTA COLLEGE'='NWV',
                        'ALAMO CCD PALO ALTO COLLEGE'='PAC',
                        'ALAMO CCD SAN ANTONIO COLLEGE'='SAC',
                        'ALAMO CCD ST. PHILIPS COLLEGE'='SPC',
                        'OUR LADY OF THE LAKE UNIV/SA'='OLLU',
                        "ST. MARY'S UNIVERSITY"='StMU',
                        'TEXAS A&M UNIV-SAN ANTONIO'='TAMUSA',
                        'UNIV OF THE INCARNATE WORD'='IWC'))

Bar_Graph2 <- ggplot(data=Outcome, aes(x = College, y=Grads_From_NISD, fill=College)) +
  geom_bar(stat="identity") +
  #coord_flip() + 
  theme_classic() +
  theme(axis.text.y = element_text(face="bold")) + 
  scale_fill_manual(values = mycolors) +
  labs(y="Students  Earning  Degree") +
  theme(axis.title.x = element_blank()) +
  theme(legend.position = "none")
```

Column {data-width=400}
-------------------------------------
### **Local Higher Education**

Over the last three years, enrollment of NISD HS graduates in local higher education institutions has risen `r Pct_Change`% percent to `r sum(HS2019_sk3$LocalCollege)` students.

The interactive graphs/table to the right depict: (1) the trend in NISD graduates enrolling locally; and, (2) the number of students earning at least one degree within 6 years (Class of 2008-2010, the latest figures available).  

Key findings include:  
$*$ Northwest Vista College (NWV) comprises more than half of the total NISD graduate enrollees.  
$*$ The number of NISD graduates enrolling at UTSA has declined each of the last two years while San Antonio College (SAC) nearly doubled in 2019 enrollments over 2018 levels.  
$*$ UTSA enrollees had the highest number of degrees earned, followed closely by NWV.


Column {.tabset}
-------------------------------------
### Graph: College Trends


```{r echo = FALSE, warning = FALSE, error = FALSE}
ggplotly(Line_Graph1,tooltip = "y")
```

### Graph: College Outcomes

```{r echo = FALSE, warning = FALSE, error = FALSE}
ggplotly(Bar_Graph2,tooltip = "y")
```

